---
title: "Getting Started with Roller"
author: "Sharon Hui"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Roller}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
library(Roller)
```

## Device Rolling

The package `"Roller"` is a minimal implementation for simulating rolling a device a given number of times. `"Roller"` can also be used to visualize the relative frequencies in such a series of rolls.

## Creating a `device`

The first step would be to create a `"device"` object with the function `device()`:

```{r}
fair_coin <- device()
fair_coin

```

By default, `device()` creates a _fair_ coin with standard sides `1`, `2`, and with probability values for `1` to be 0.5 and `2` to be 0.5.

You can change the default arguments `sides` and `prob` in the `device()`.

```{r}
weird_die <- device(
  sides = c('i', 'ii', 'iii', 'iv'),
  prob = rep(1/4, 4)
  )
weird_die
```

## Rolling the device

After you have defined the `"device"` object, you can use the function `roll()` to roll the device a specified number of `times`. The output of `roll()` will be a list with the following four elements:

 - `rolls`: vector with outputs of the rolls
 - `sides`: vector with the sides of the `"device"` object
 - `prob`: vector with probabilities for each side of the `"device"` object
 - `total`: total number of rolls (i.e. `times`)
 
However, `roll()` will only print the class of the object and the generated `rolls`. In other words, it will display the class of the object (`rolls`) and the rolls. To call/display the other elements such as `sides`, `prob`, and `total`, use `$`.

```{r}
# create a device called fair_die with sides, 1 to 6, each with a probability of 1/6
fair_die <- device(sides = 1:6, prob = rep(1/6, 6))
# roll fair_die 50 times
set.seed(123)
fair_50rolls <- roll(fair_die, times = 50)

# print the object
fair_50rolls

# to display the rolls vector with outputs of the rolls
fair_50rolls$rolls

# to display the sides vector with the sides of the `"device"` object
fair_50rolls$sides
  
# to display the prob vector with probabilities for each side of the `"device"` object
fair_50rolls$prob

# to display the total number of rolls
fair_50rolls$total

# to display what is in fair_50rolls
names(fair_50rolls)
```

Here is another example of using the rolls.

```{r}
# creating a string die device
str_die <- device(
sides = c('a', 'b', 'c', 'd', 'e', 'f'),
prob = c(0.075, 0.1, 0.125, 0.15, 0.20, 0.35))
# roll str_die 20 times
set.seed(123)
str_rolls <- roll(str_die, times = 20)

# print the object
str_rolls

# to display the rolls vector with outputs of the rolls
str_rolls$rolls

# to display the sides vector with the sides of the `"device"` object
str_rolls$sides
  
# to display the prob vector with probabilities for each side of the `"device"` object
str_rolls$prob

# to display the total number of rolls
str_rolls$total


# to display what is in fair_50rolls
names(str_rolls)

```

The `summary()` function for rolls also gives you a table of frequencies including the side (`side`), count (`count`), and proportion (`prop`). The summary method returns an object, `"summary.rolls"` and the otput wil be a list containing a data frame called `freqs` with the three columns, `side`, `count`, and `prop`.

 - The `side` column gives the sides of the rolled `device`.
 - The `count` column gives the frequency (count) of each side of the rolled device.
 - The `prop` column gives the relative frequency (proportion) of each side of the rolled `device`.


```{r}
# creates the summary.rolls object for str_rolls
summary(str_rolls)

# to display the side column of the `device`
summary(str_rolls)$freq$side

# to display the count column of the `device`
summary(str_rolls)$freq$count

# to display the prop column of the `device`
summary(str_rolls)$freq$prop

# to display the names of the `summary.rolls` device, `str_rolls`
names(summary(str_rolls))

# to display the class of str_rolls
class(summary(str_rolls))
```

## More Functions for the Object `"rolls"`

You can use the `"["` to extract the value of the given roll. This is an extraction method.
```{r}
# roll fair 8-sided die
set.seed(123)
fair_dev <- device(sides = letters[1:8], prob = rep(1/8, 8))
fair500 <- roll(fair_dev, times = 500)
# summary method
summary(fair500)

# extracting the roll in position 500
fair500[500]

```

You can use the `"[<-"`to replace the value of a given roll. This is an replacement method.

```{r}
# replacing a roll in position 500
fair500[500] <- 'a'
fair500[500]

```

This will change the summary of `fair500`.
```{r}
# look at the new summary for fair 500
summary(fair500)
```

You can use the `"+"` to add more rolls. This is an addition method.
```{r}
# adding 100 rolls
fair600 <- fair500 + 100

# this is the summary of fair600
summary(fair600)
```


## Plotting Counts and Relative Frequencies of Rolls

You can use the function `plot()` to create a graphic with the relative frequencies of a series of device rolls:

```{r, fig.show='hold', fig.width = 7, fig.height = 4, fair_50rolls}
# plotting fair_50rolls using the plot method
plot(fair_50rolls)
```
```{r, fig.show='hold', fig.width = 7, fig.height = 4, fair500}
# plotting fair500 using the plot method
plot(fair500, 500)
```


























