---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```


## Overview

`"Roller"` is a minimal [R](http://www.r-project.org/) package that provides functions that simulate rolling a device.

* `device()` creates a device object (of class `"device"`)
* `roll()` rolls a device object, producing a `"rolls"` object.
* `plot()` method for a `"rolls"` object to plot the relative frequencies for each side in a simulation.
* `summary()` method for a `"rolls"` object which outputs the class of the object (`summary.rolls`) and a data frame with the `side`, `count`, and `prop`.


## Motivation

This package was created to illustrate how to create a basic R package and to demonstrate core concepts on how an R package works.


## Installation

Install the development version from GitHub via the package `"devtools"`:

```r
# development version from GitHub:
#install.packages("devtools") 
# install "Roller" (without vignettes)
devtools::install_github("hw-stat133-swhui/Roller")
# install "Roller" (with vignettes)
devtools::install_github("hw-stat133-swhui/Roller", build_vignettes = TRUE)
```


## Usage

```{r}
library(Roller)
# default device
fair_coin <- device()
fair_coin

# create a device called fair_die with sides, 1 to 6, each with a probability of 1/6
fair_die <- device(sides = 1:6, prob = rep(1/6, 6))
# roll fair_die 50 times
set.seed(123)
fair_50rolls <- roll(fair_die, times = 50)

# print the object
fair_50rolls

# to display the rolls vector with outputs of the rolls
fair_50rolls$rolls

# to display the sides vector with the sides of the `"device"` object
fair_50rolls$sides
  
# to display the prob vector with probabilities for each side of the `"device"` object
fair_50rolls$prob

# to display the total number of rolls
fair_50rolls$total

# to display what is in fair_50rolls
names(fair_50rolls)

# creates and displays the summary.rolls object for fair_50rolls
obj <- summary(fair_50rolls)

# to display the side column of the `device`
summary(fair_50rolls)$freq$side
# or
obj$freq$side

# to display the count column of the `device`
summary(fair_50rolls)$freq$count

# to display the prop column of the `device`
summary(fair_50rolls)$freq$prop

# to display the names of the `summary.rolls` device, `fair_50rolls`
names(summary(fair_50rolls))

# to display the class of fair_50rolls
class(summary(fair_50rolls))

# extraction method for fair_50rolls
fair_50rolls[1]

# replacement method for fair_50rolls
fair_50rolls[1] <- 2
fair_50rolls[1]

# look at the new summary for fair_50rolls
summary(fair_50rolls)

# adding more rolls to fair_50rolls
fair_70rolls <- fair_50rolls + 20

# display summary of fair_70rolls
summary(fair_70rolls)
```


```{r Relative_freq_50_rolls}
# plotting fair_50rolls using the plot method
plot(fair_50rolls)
```
